---
title: "DoSomething Data Analyst Intern Assignment"
date: "March 16, 2018"
author: "Michelle Chung"
output:
  html_notebook:
    theme: cosmo
    
---

<hr>

This document explores the DoSomething dataset "member_event_log_20180305.csv". It is intended to provide a brief summary and present initial findings of the data based the DoSomething Data Analyst intern assignment guidelines.

I make the following assumptions about the product in my analysis:
<ul>
<li>In the life cycle of the user, account creation occur only once per unique user.</li>
<li>A sign-up indicates that a user has signed up to participate in a task or an event.</li>
<li>A reportback indicates that a user has marked that they have completed a task or event.</li>
</ul>

<h3>Summary</h3>

```{r}
my_data <- read.csv(file="member_event_log_20180305.csv", header=TRUE, sep=",")
data_summary <- summary(my_data)
data_summary
```

The data_summary tells us that this dataset records a unique event identified by an `event_id`. The `user_id`, `action_type`, `channel`, and `timestamp` are recorded for each event.

<h3>Analysis</h3>

Let's go for the low-hanging fruit. A summary glance at the data can tell us the following:

<h4>The majority of user actions occur on web.</h4>

```{r, echo=FALSE}
channels <- table(my_data$channel)
total <- sum(channels)
p_web = channels["web"]/total
p_paid = channels["paid"]/total
p_sms = channels["sms"]/total
```

But not by a lot. `r floor(channels["web"]/total*100)`% of actions occur on web, compared to `r floor(channels["paid"]/total*100)`% in paid and `r floor(channels["sms"]/total*100)`% in SMS.

<h4>There is a clear discrepancy in the amount of sign-ups and the amount of report-backs.</h4>
<br>

```{r, echo=FALSE}
action_types <-table(my_data$action_type)
y_axis = names(action_types)
dotchart(as.numeric(action_types),y_axis,
  	main="User Action Types", 
    xlab="Number of Events")
```

It appears that sign-ups are the most common user actions--and that's great! But reportbacks give us a better understanding of how many users are actually participating in these events. Perhaps we could add or alter a feature that makes reportbacks easier to complete.

<h3>Notes</h3>

Given more time, I would have liked to explore the following:

<ul>
<li>The distribution of events over time. This was the big one that I really wanted to do, but wasn't able to get to complete cohesively under the four-hour time limit. My initial approach was to go from really large (month-by-month) to really granular (minute-by-minute). The time data was, in my opinion, the juiciest feature--perhaps I should have tackled it first. I think predictive models based on the time data would have helped support strong product recommendations.</li>

<li>To add to the above point: Distribution of the different action types over time. At what times are people most likely to perform certain actions?</li>

<li>Sort the data by least to most active users (based on number of actions). Who are the power users? What metric can we produce, other than collecting events, that can tell us about user engagement?</li>

<li>How can we use the data to understand what makes someone likely to report back? Are there other events that correlate with a reportback?</li>

</ul>

In regards to enriching the data, I would have loved to see more information about the users as well as the events they signed up for. For users, features like age, occupation, and location would have been interesting to explore; for events, time/date data, location, event organizers, and information about event attendees are just a few things I can think of that would have yielded more interesting results.